Developing a preprint culture in biology

### 2013-02-27 ms.tex
edit Ethan's edits :P

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new commercial open
access publisher that provides a preprint server and a peer reviewed journal. It
is focused on the biological sciences, which may counter the perception that
preprints do not have a place in biology. Preprints can optionally be made
private.  One preprint per year can be posted for free, with a onetime
(\emph{i.e.}  lifetime) fee for unlimited public preprints. Preprints can be
posted to PeerJ regardless of where they will be submitted for publication.

@@

The ongoing discussions on the scientific publication process, peer-review and
alternative publication models are all symptoms of the current uneasiness with
the ever growing obsession with bibliographic metrics such as the impact factor
\cite{Fisher2012}. Researchers are pressured to orient their publication
strategy to maximize their number of publications and total citations. A
well-known consequence is the initial submission of manuscripts to the most
prestigious journals, followed by resubmission to progressively ``lower level''
journals as they are rejected. The numerous negative impacts of such behavior
have been discussed in depth \cite{hoc09} and include a long delay between the
time a manuscript is finished and its publication. Research activities and the
publication process are drifting away from their fundamental objective, namely
the distribution of novel scientific discoveries.

Developing a preprint culture in biology will not solve all of the problems with
the current publication process. However, it might significantly reduce the
negative consequences. The role of peer-review is to judge the scientific
quality of a study. It is the first barrier against the fraudulent and poor
quality science that could impede scientific progress. In practice, the
peer-review system is not only used to evaluate scientific quality but also to
judge pertinence. On the other hand, preprints are not filtered on either
quality or pertinence. Widespread adoption of preprint servers has the potential
to shift the distribution strategy: journals would remain important to validate
publications, but the relevance of a study should only be judged by many more
readers than the typical two to four anonymous reviewers. With a shift in the
diffusion strategy, the role of traditional journals and their editors would be
to showcase scientific discoveries for specialized readership.

Making publication easier can lead to the proliferation of studies of uneven 
quality. A trade-off between the intensity of the peer-review filtering and 
the benefits to science has been hypothesized \cite{Aarssen2012}.  With 
increasingly stringent peer review, the quality of published papers can 
improve at the cost of an increased load on authors and 
reviewers, and greater delays for publication.  By
bypassing this model preprints 
speed up the dissemination of scientific discoveries and put on reader's 
shoulders the responsibility to judge originality and pertinence.

@@

DG is an associate falculty of F1000, EW is an academic editor for PeerJ, and TP
is an advisor for figshare.

>>>>

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new commercial open
access publisher focused on the biological sciences that provides a preprint
server and a peer reviewed journal. Preprints can optionally be made private.
One preprint per year can be posted for free, with a onetime (\emph{i.e.}
lifetime) fee for unlimited public preprints. Preprints can be posted to PeerJ
regardless of where they will be submitted for publication.

@@

The ongoing discussions on the publication process, peer-reviewing and 
alternative publication models are all symptoms of the current uneasiness 
with the ever growing obsession with bibliographic metrics such as the 
impact factor \cite{Fisher2012}. Researchers are pressured to orient their 
publication strategy to maximize their number of publications and total 
citations. A well-known consequence is to submit manuscripts first to the 
most prestigious journals, and then resubmit to ``lower level'' journals as 
they are rejected. The numerous negative impacts of such behavior have been 
discussed in depth \cite{hoc09} and include a long delay between the time a 
manuscript is finished to its publication. Research activities and the 
publication process are drifting away from their fundamental object, namely 
the diffusion of novel scientific discoveries.

Developing a preprint culture in biology will not solve all problems with 
the current publication process. However, it might significantly reduce its 
negative consequences. The role of peer-reviewing is to judge the scientific 
quality of a study. It is the first barrier against the fraudulent and poor 
quality science susceptible to impede scientific progress. In practice, the 
peer-review system is not only used to evaluate scientific quality but also 
to judge pertinence. On the other hand, preprints are not filtered, neither 
for their quality nor their pertinence. Widespread adoption of preprint 
servers has the potential to shift the diffusion strategy: journals would 
remain important to validate publications, but the relevance of a study 
should only be judged by many more readers than the typical two-four 
anonymous reviewers. With a shift in the diffusion strategy, the role of 
traditional journals and their editors would be to showcase scientific 
discoveries for specialized readership.

Making publication easier can lead to the proliferation of studies of uneven 
quality. A trade-off between the intensity of the peer-review filtering and 
the benefits to science has been hypothesized \cite{Aarssen2012}.  With 
increasingly stringent peer reviewing, the quality of published papers can 
improve at the cost of an increased load on authors and 
reviewers and greater delays for publication.  Preprints are simply 
bypassing this model for what we believe is the progress of science: they 
speed up the dissemination of scientific discoveries and put on reader's 
shoulders the responsibility to judge originality and pertinence.

@@

EW is an academic editor for PeerJ and TP is a member of the figshare advisors
program.
